---
title: "Data visualization"
author: "Duzhe Wang"
date: "8/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Library

```{r}
library(tidyverse)
```

## The ```mpg``` data frame

```{r}
mpg
summary(mpg)
```


## Creating a ggplot 

```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=hwy))


```


- ```ggplot()```: create a coordinate system that you can add layers to
- ```geom_point()```: add a layer of points to the plot
- Each geom function in ggplot2 takes a mapping argument, which defines how variables in the dataset are mapped to visual properties. 

## A graphing template 

```ggplot(date=<DATA>)+<GEROM_FUNCTION>(mapping=aes(<MAPPINGS>))```



## Aesthetic mappings

- To map an aesthetic to a variable, associate the name of the aesthetic to the name of the variable inside ```aes()```. 
ggplot2 will automatically assign a unique level of the aesthetic to each unique value of the variable, a process known as scaling. 
ggplot2 will also add a legend that explains which levels correspond to which values. 


- color

```{r}

ggplot(data=mpg)+geom_point(map=aes(x=displ, y=hwy, color=class))


```

- alpha: control the transparency 

```{r}

ggplot(data=mpg)+geom_point(map=aes(x=displ, y=hwy, alpha=class))

```

- shape: control the shape

```{r}
ggplot(data=mpg)+geom_point(mapping=aes(x=displ, y=hwy, shape=class))

```

- We can also set the aesthetic properties of the geom manually. For example, 

```{r}
ggplot(data=mpg)+geom_point(map=aes(x=displ, y=hwy), color="blue")

```

- stroke: [control the thickness of point boundary](https://stagraph.com/HowTo/Plot/Aes/Stroke)

```{r}
ggplot(data=mpg)+geom_point(mapping=aes(x=displ, y=hwy), size=5, color="white", stroke=1)

ggplot(data=mpg)+geom_point(mapping=aes(x=displ, y=hwy), size=5, color="white", stroke=2)
```


- **Common probnlem**: the ```+``` sign has to come at the end of the line, not the start. 


## Facets

- One way to add additional variables is with aesthetics. Another way, particularly useful for 
categprical variables, is to split your plot into facets, subplots that each display one subset of the data. 

- To facet your plot by a single variable, use ```facet_wrap()```. The first argument of ```facet_wrap()``` should be a formula, which you create with ```~``` followed by a variable name. The variabe that you pass to ```facet_wrap()```should be discrete. 

```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=hwy))+
  facet_wrap(~class, nrow=2)
```


- To facet your plot on the combination of two variables, add ```facet_grid()``` to your plot call. The first argument of ```facet_grid()``` is also a formula. This time the formula should contain two variable names separated by a ```~```. 

```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=hwy))+
  facet_grid(drv~cyl)
```

- If you prefer to not facet in the rows or columns dimension, use ```.``` instead of a variable name. 

```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=hwy))+
  facet_grid(.~cyl)

```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)
```

## Geometric objects 

- A geom is the geometrical object that a plot uses to represent data. 

```{r}
gg=ggplot(data=mpg)

gg+geom_point(mapping=aes(x=displ, y=hwy))

gg+geom_smooth(mapping=aes(x=displ, y=hwy))

```



```{r}
gg+geom_smooth(mapping=aes(x=displ, y=hwy, linetype=drv))

```


```{r}
gg+
  geom_point(mapping=aes(x=displ, y=hwy, color=drv))+
  geom_smooth(mapping=aes(x=displ, y=hwy, linetype=drv))
```


```{r}
ggplot(data=mpg, mapping=aes(x=displ, y=hwy))+
  geom_point()+
  geom_smooth()
```






```{r}
gg+geom_smooth(mapping=aes(x=displ, y=hwy, group=drv))
```

```{r}
gg+geom_smooth(mapping=aes(x=displ, y=hwy, color=drv), show.legend=FALSE)


```

```{r}
ggplot(data=mpg, mapping=aes(x=displ, y=hwy))+
  geom_point(mapping=aes(color=class))+
  geom_smooth()

```



```{r}
ggplot(data=mpg, mapping=aes(x=displ, y=hwy))+
  geom_point(mapping=aes(color=class))+
  geom_smooth(data=filter(mpg, class=="subcompact"), se=FALSE)


```


```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```


- Exercise 6 

```{r}
ggplot(data=mpg, mapping=aes(x=displ, y=hwy))+
  geom_point()+
  geom_smooth(color="blue", se=FALSE)

```

```{r}
ggplot(data=mpg, mapping=aes(x=displ, y=hwy))+
  geom_point()+
  geom_smooth(aes(group=drv), se=FALSE)
```


```{r}
ggplot(data=mpg, mapping=aes(x=displ, y=hwy, color=drv))+
  geom_point()+
  geom_smooth(se=FALSE)
```


```{r}
ggplot(data=mpg )+
  geom_point(mapping=aes(x=displ, y=hwy, color=drv))+
  geom_smooth(mapping=aes(x=displ, y=hwy), se=FALSE)

```


```{r}
ggplot(data=mpg, mapping=aes(x=displ, y=hwy))+
  geom_point(mapping=aes(color=drv))+
  geom_smooth(aes(linetype=drv), se=FALSE)


```


```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=hwy, color=drv), stroke=2 )

```


## Statistical transformations

- bar chart

```{r}
diamonds
```

```{r}
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut))
```

- You can recreate the previous plot using ```stat_count()``` instead of ```geom_bar()```:

```{r}
ggplot(data=diamonds)+
  stat_count(mapping=aes(x=cut))
```


- Three reasons to use a stat explicitly: 

```{r}
demo=tribble(
  ~cut, ~freq,
  "Fair", 1610,
  "Good", 4906,
  "Very good", 12082,
  "Premium", 13791,
  "Ideal", 21551
)



ggplot(data=demo)+
  geom_bar(mapping=aes(x=cut, y=freq), stat="identity")
```


```{r}
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut, y=stat(prop), group=1))



```


```{r}
ggplot(data=diamonds)+
  stat_summary(
    mapping=aes(x=cut, y=depth), 
    fun.ymin=min,
    fun.ymax=max,
    fun.y=median
  )


```

## position adjustments 


```{r}
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut, color=cut))


ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut, fill=cut))


```















